Social Participation Ontology: community documentation, enhancements and use examples

### 2015-01-08 paper.tex
adiantado script para sintese da OPS e adicionada figura elaborada da OPS
atualizando artigo com boa parte do que faltava
finalmente terminada introducao

\preprint{XXXXX (preprint)}

\title[Ontology of Social Participation (OSP)]{Ontology of Social Participation: Community Documentation, Expansion and Examples of Use}% Force line breaks with \\

@@

\begin{itemize}
    \item Various proposals for the OSP logo, some of which are in figure~\ref{logo}.
    \item Figure~\ref{fig:v1} shows an English and completed version of the original diagram of CVSP in the document~\cite{conceptualMod}.
    \item A diagram for general public consultations. An OWL implementation of this model is envisioned for near future, as described in subsection~\ref{public}. Figure~\ref{fig:consult} has the diagram itself.
\end{itemize}

@@

\subsection{Discussions and etherpads}
There are numerous commentaries in the blog posts, in which collective elaboration was registered. Besides that, four etherpads 
were written (these are interfaces that allows writting online texts with multiple simulteneous contributors~\cite{etherpads}):
\begin{itemize}
    \item A pad for important words.
    \item A pad dedicated to a second phase of the common vocabulary elaboration, which did not happen yet.
    \item A pad for process documentation. It became the first document described in Subsection~\ref{refDocs}. 
    \item A pad for both vocabulary specification and ``questions not addressed to in the webinar''.
\end{itemize}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/opsBFO}
    \caption{Diagram representation of the Ontology of Social Participation (OSP). Arrows with white heads indicates ``is a'' relations (subclass points to superclass). Arrows with black heads indicates property relations from subject to object . All property relation yield existential restrictions, with the exception of the ``has characteristic'' property, that does not yield restriction.}
    \label{fig:v1}
\end{figure*}

Followng the exposition in Section~\ref{sec:into}, from the CVSP the OSP was conceived. This section makes four considerations about OSP: first is label standardization and an exhibition of implemented and unimplemented CVSP classes, properties and restrictions. Features present in Figure~\ref{fig:v1}, but not present in the community release, are fully implemented in Section~\ref{impl}. Elementary further specification of subclasses is addressed in Section~\ref{downwards}. Examples of usage are addressed in subsection~\ref{examplesUsage}.

For completeness of exposition and deference to CVSP community efforts, consideration of upper ontologies is at Subsection~\ref{upper} and the public consultation model depicted in Figure~\ref{fig:consult} is considered at Subsection~\ref{public}.

@@

\lstinputlisting[language=owlms]{OWL/cospMan.owl}
\nocite{*}
\bibliography{paper}% Produces the bibliography via BibTeX.

>>>>

%\preprint{XXXXX (preprint)}

\title[OPS: Social Participation Ontology]{Social Participation Ontology: community documentation, enhancements and use examples}% Force line breaks with \\

@@

\begin{itemize}
    \item Various proposals for the OSP logo, some of which are in figure~\ref{logo}.
    \item Figure~\ref{fig:v1} shows an English and completed version of the original diagram of CVSP in the document~\cite{conceptualMod}.
    \item A diagram for general public consultations. An OWL implementation of this model is envisioned for near future, as described in Section~\ref{public}. Figure~\ref{fig:consult} has the diagram itself.
\end{itemize}

@@

\subsection{Discussions and etherpads}
There are numerous commentaries in the blog posts, in which collective elaboration was registered. Besides that, four etherpads 
were written (these are interfaces that allows writting online texts with multiple simulteneous contributors~\cite{etherpads}):
\begin{itemize}
    \item A pad for important words.
    \item A pad dedicated to a second phase of the common vocabulary elaboration, which did not happen yet.
    \item A pad for process documentation. It became the first document described in Section~\ref{refDocs}. 
    \item A pad for both vocabulary specification and ``questions not addressed to in the webinar''.
\end{itemize}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/opsBFO___}
    \caption{Diagram representation of the Ontology of Social Participation (OSP). Arrows with white heads indicates ``is a'' relations (subclass points to superclass). Arrows with black heads indicates property relations from subject to object . All property relation yield existential restrictions, with the exception of the ``has characteristic'' property, that does not yield restriction.}
    \label{fig:v1}
\end{figure*}

Followng the exposition in Section~\ref{sec:into}, from the CVSP the OSP was conceived. This section makes four considerations about OSP: first is label standardization and an exhibition of implemented and unimplemented CVSP classes, properties and restrictions. Features present in Figure~\ref{fig:v1}, but not present in the community release, are fully implemented in Section~\ref{impl}. Elementary further specification of subclasses is addressed in Section~\ref{downwards}. Examples of usage are addressed in Section~\ref{examplesUsage}.

For completeness of exposition and deference to CVSP community efforts, consideration of upper ontologies is at Section~\ref{upper} and the public consultation model depicted in Figure~\ref{fig:consult} is considered at Section~\ref{public}.

@@

\lstinputlisting[language=owlms]{rdf/ops.ttl}
\nocite{*}
\bibliography{paper}% Produces the bibliography via BibTeX.
